Subject-specific curves for time series forecasting of smart meter demand

### 2019-11-13 sscts-paper-1.Rmd
Update feature selection plot to colour scaled_temperature and summary statistics differently.

```{r fs-alpha, fig.cap="(ref:fs-alpha)", fig.height=7*1.4}
fs_df %>%
  select(period, month, vars) %>%
  unnest() %>%
  mutate(
    selected = TRUE,
    month = month(month, label = TRUE)
  ) %>%
  spread(vars, selected, fill = FALSE) %>%
  # Ensure models with no variables selected have white squares plotted
  right_join(list(
    month = unique(month(fcst_test_dates)),
    period = fcst_test_periods
  ) %>%
    cross_df() %>%
    mutate(month = month(month, label = TRUE)),
  by = c("month", "period")
  ) %>%
  gather(vars, selected, -c(period, month)) %>%
  mutate(
    selected = tidyr::replace_na(selected, FALSE),
    vars = fct_rev(factor(vars, levels = predictor_vars)),
    season = if_else(month %in% month(c(10:12, 1:3), label = TRUE),
      "October-March", "April-September"
    )
  ) %>%
  group_by(season, period, vars) %>%
  summarise(n_selected = sum(selected)) %>%
  ungroup() %>%
  mutate(
    selected = if_else(n_selected > 0, TRUE, FALSE),
    predictor_type = case_when(
      str_detect(vars, "^scaled_temperature") ~ "Temperature",
      str_detect(vars, "^scaled_wh") ~ "Demand",
      TRUE ~ "Other"
    )
  ) %>%
  filter(n_selected > 0) %>% # don't want any fill for 0
  ggplot(aes(
    x = period, y = vars, alpha = n_selected,
    fill = predictor_type
  )) +
  geom_tile() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  ) +
  labs(
    x = "Period",
    y = "Predictor",
    fill = "Predictor type",
    alpha = "Number of times selected"
  ) +
  facet_wrap(~season, ncol = 1) +
  scale_x_continuous(breaks = seq(8, 96, 8))
```

@@

```{r fs-all-periods-months, fig.cap="(ref:fs-all-periods-months)", fig.height = 7*1.4}
fs_df %>%
  select(period, month, vars) %>%
  unnest() %>%
  mutate(
    selected = TRUE,
    month = month(month, label = TRUE)
  ) %>%
  spread(vars, selected, fill = FALSE) %>%
  # Ensure models with no variables selected have white squares plotted
  right_join(list(
    month = unique(month(fcst_test_dates)),
    period = fcst_test_periods
  ) %>%
    cross_df() %>%
    mutate(month = month(month, label = TRUE)),
  by = c("month", "period")
  ) %>%
  gather(vars, selected, -c(period, month)) %>%
  mutate(
    selected = tidyr::replace_na(selected, FALSE),
    vars = fct_rev(factor(vars, levels = predictor_vars)),
    predictor_type = case_when(
      str_detect(vars, "^scaled_temperature") ~ "Temperature",
      str_detect(vars, "^scaled_wh") ~ "Demand",
      TRUE ~ "Other"
    )
  ) %>%
  filter(selected == TRUE) %>%
  ggplot(aes(x = period, y = vars, fill = predictor_type)) +
  geom_tile() +
  facet_wrap(~month) +
  # scale_fill_manual(values = c("white", "#3d6b78")) +
  labs(
    x = "Period",
    y = "Predictor",
    fill = "Predictor type"
  ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(8, 96, 8))
```
\newpage

>>>>

```{r fs-alpha, fig.cap="(ref:fs-alpha)", fig.height=7*1.4}
fs_df %>%
  select(period, month, vars) %>%
  unnest() %>%
  mutate(
    selected = TRUE,
    month = month(month, label = TRUE)
  ) %>%
  spread(vars, selected, fill = FALSE) %>%
  # Ensure models with no variables selected have white squares plotted
  right_join(list(
    month = unique(month(fcst_test_dates)),
    period = fcst_test_periods
  ) %>%
    cross_df() %>%
    mutate(month = month(month, label = TRUE)),
  by = c("month", "period")
  ) %>%
  gather(vars, selected, -c(period, month)) %>%
  mutate(
    selected = tidyr::replace_na(selected, FALSE),
    vars = fct_rev(factor(vars, levels = predictor_vars)),
    season = if_else(month %in% month(c(10:12, 1:3), label = TRUE),
      "October-March", "April-September"
    )
  ) %>%
  group_by(season, period, vars) %>%
  summarise(n_selected = sum(selected)) %>%
  ungroup() %>%
  mutate(
    selected = if_else(n_selected > 0, TRUE, FALSE),
    predictor_type = case_when(
      str_detect(vars, "^scaled_temperature$") ~ "Forced variable",
      str_detect(vars, "^scaled_temperature_lag") ~ "Temperature lags",
      str_detect(vars, "^scaled_temperature_(?!lag)") ~ "Temperature statistic",
      str_detect(vars, "^scaled_wh_lag") ~ "Demand lags",
      TRUE ~ "Other"
    )
  ) %>%
  filter(n_selected > 0) %>% # don't want any fill for 0
  ggplot(aes(
    x = period, y = vars, alpha = n_selected,
    fill = predictor_type
  )) +
  geom_tile() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  ) +
  labs(
    x = "Period",
    y = "Predictor",
    fill = "Predictor type",
    alpha = "Number of times selected"
  ) +
  facet_wrap(~season, ncol = 1) +
  scale_x_continuous(breaks = seq(8, 96, 8))
```

@@

```{r fs-all-periods-months, fig.cap="(ref:fs-all-periods-months)", fig.height = 7*1.4}
fs_df %>%
  select(period, month, vars) %>%
  unnest() %>%
  mutate(
    selected = TRUE,
    month = month(month, label = TRUE)
  ) %>%
  spread(vars, selected, fill = FALSE) %>%
  # Ensure models with no variables selected have white squares plotted
  right_join(list(
    month = unique(month(fcst_test_dates)),
    period = fcst_test_periods
  ) %>%
    cross_df() %>%
    mutate(month = month(month, label = TRUE)),
  by = c("month", "period")
  ) %>%
  gather(vars, selected, -c(period, month)) %>%
  mutate(
    selected = tidyr::replace_na(selected, FALSE),
    vars = fct_rev(factor(vars, levels = predictor_vars)),
    predictor_type = case_when(
      str_detect(vars, "^scaled_temperature$") ~ "Forced variable",
      str_detect(vars, "^scaled_temperature_lag") ~ "Temperature lags",
      str_detect(vars, "^scaled_temperature_(?!lag)") ~ "Temperature statistic",
      str_detect(vars, "^scaled_wh_lag") ~ "Demand lags",
      TRUE ~ "Other"
    )
  ) %>%
  filter(selected == TRUE) %>%
  ggplot(aes(x = period, y = vars, fill = predictor_type)) +
  geom_tile() +
  facet_wrap(~month) +
  # scale_fill_manual(values = c("white", "#3d6b78")) +
  labs(
    x = "Period",
    y = "Predictor",
    fill = "Predictor type"
  ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(8, 96, 8))
```
\newpage
